---
title: "R Notebook"
output: html_notebook
---

Purpose: analyzing indicators related to energy with the GDP of a country.

Insert datasets and set variables
```{r}
setwd("C:/Users/lkim0/Desktop/capstone")
# load libraries
library(readxl)
library(mice)
library(dplyr)
library(stringr)
library(tidyr)
library(data.table)
library(maps)
library(ggplot2)
library(googleVis)
library(shinydashboard)
library(shiny)
library(VIM)
library(car)
library(MASS)
library(DMwR)
library(DT)

select = dplyr::select

wdi = read_excel("W03b_wdi.xlsx")
y = "NY.GDP.MKTP.CD" # (GDP (current US$)
```

Cleaning and manipulating the data:
# filter unnecessary countries: Early-demographic dividend, (excluding high income), (IDA & IBRD countries), IDA, Heavily indebted poor countries (HIPC), High income, Late-demographic     dividend, Least developed countries: UN classification, Low & middle income, Low income, Lower middle income, OECD members, Other small states, Post-demographic dividend, Pre-demographic dividend, Small states, Upper middle income, World, Region country code
```{r}
# renaming wdi
qual = c("country.name", "country.code", "indicator.name", "indicator.code")
yr = 1960:2017
yr = paste("x", yr, sep = "")
colnames(wdi) = c(qual, yr)
wdi$indicator.name = sapply(wdi$indicator.name, tolower)

rm.reg = c("ARB","CEB","CSS","EAP","EAR","EAS","ECA","ECS","EUU","HIC","HPC","IBD","IBT","IDA","IDX",
           "LAC","LCN","LDC","LIC","LMC","LMY","MEA","MIC","MNA","NAC","OED","OSS","PRE","PSS","PST",
           "SAS","SSA","SSF","SST","TEA","TEC","TLA","TMN","TSA","UMC","WLD", "FCS", "INX", "IDB", "LTE")

# Remove those regions, non-country data rows
wdi = wdi %>% filter(!country.code %in% rm.reg)
# filter out indicator codes related to energy and GDP
en.word = c("energy", "electricity", "clean", "emission", "electrical", "power")
ind.df = wdi %>% filter(str_detect(indicator.name, en.word[1]))

  for(a in 2:length(en.word)) {
    ind.df = rbind(ind.df, wdi %>% filter(str_detect(indicator.name, en.word[a])))
  }

ind.filter = as.data.frame(table(ind.df$indicator.name))

final.df = wdi %>% filter(indicator.name %in% ind.filter$Var1 | indicator.code %in% y) %>%
  select(one_of(qual, "x2015"))

final.df = dcast(melt(as.data.table(final.df), id.vars = c("country.name", 
                                                        "country.code","indicator.name", "indicator.code")), 
               country.name + country.code + variable ~ indicator.code, value.var = "value")

final.df = final.df %>% select(-variable)
head(final.df)

i = colnames(comp.df)
ind.df = wdi %>% filter(country.code %in% "USA") %>% filter(indicator.code %in% i) %>% select(one_of(c("indicator.name","indicator.code")))
ind.df

```

Deal with missingness:
```{r}
na = final.df %>% summarize_all(funs(sum(is.na(.))/n()))
na = gather(na, key="feature", value="missing_pct")
goodv = filter(na, missing_pct < 0.25)
final.df = final.df %>% select(one_of(c(goodv$feature))) # filter NA
# note: make sure to keep response variable after cleaning out the NAs
final.df$variable = as.numeric(gsub("x", "", final.df$variable)) # change variable to a number
#final.df$NY.GDP.MKTP.CD = final.df$NY.GDP.MKTP.CD/1000000000000 # GDP in the trillions

summary(final.df)
```

analyze the missingness
```{r}
aft.transpose = c("country.name","country.code","variable")
# visual analysis of missing data using mice functions
par("mar")
par(mar = c(1,1,1,1))
md.pattern(final.df)

mice.plot <- aggr(final.df, col=c('navyblue','yellow'),
                  numbers=TRUE, sortVars=TRUE,
                  labels=names(final.df), cex.axis=.7,
                  gap=3, ylab=c("Missing data","Pattern"))
mice.plot
```

impute using mice
- there are 16 datasets that are created through this imputation method
```{r}
# clean the data for imputation
miss.df = final.df %>% select(-one_of(aft.transpose))
miss.df = scale(miss.df) # scale the data to impute it

imputed.data <- mice(miss.df, m = 16, maxit = 50, method = 'pmm', seed = 500, trace = F)
summary(imputed.data)

comp.df <- mice::complete(imputed.data,8)
comp.df

i = colnames(comp.df)
ind.df = wdi %>% filter(country.code %in% "USA") %>% filter(indicator.code %in% i) %>% select(one_of(c("indicator.name","indicator.code")))
ind.df
```

Training the model:
```{r}
### split train and test set
smp_size = floor(0.75*nrow(comp.df))
set.seed(123)
train.i = sample(seq_len(nrow(comp.df)), size = smp_size)
train = comp.df[train.i,]
test = comp.df[-train.i,]


### testing to find the best model ###
# note: the higher the VIF the higher the correlation
fit = lm(NY.GDP.MKTP.CD ~ ., data = train)
summary(fit)
vif(fit) # VIF (EG.ELC.ACCS.ZS, EG.ELC.ACCS.RU.ZS, **EG.ELC.ACCS.UR.ZS)
```

model 2
```{r}
fit.1 = lm(NY.GDP.MKTP.CD ~ .-EG.ELC.ACCS.ZS, data = train)
summary(fit.1)
vif(fit.1) # one of the best models / EG.CFT.ACCS.ZS & NY.ADJ.DPEM.CD are the most significant
```

model 4
```{r}
fit.4 = lm(NY.GDP.MKTP.CD ~ .-EG.ELC.ACCS.ZS-EG.ELC.ACCS.RU.ZS-EG.CFT.ACCS.ZS, data = train)
summary(fit.4)
vif(fit.4) #NOT A GOOD MODEL, LOW ADJ. R-SQUARED
```


Second training using stepAIC: looking for best model with AIC
# note: the lower the AIC the better
```{r}
fit.AIC = lm(NY.GDP.MKTP.CD ~ ., data = train)
step = stepAIC(fit, trace = FALSE)
step$anova

  # AIC final model (THIS IS THE BETTER OF THE 2 TESTING)
model = lm(NY.GDP.MKTP.CD ~ EG.CFT.ACCS.ZS + EG.ELC.ACCS.RU.ZS + NY.ADJ.DNGY.CD + NY.ADJ.DPEM.CD, data = train)
summary(model)
```

```{r}
# check what the explanatory variables are
exp.var = c("EG.CFT.ACCS.ZS", "EG.ELC.ACCS.RU.ZS", "NY.ADJ.DNGY.CD", "NY.ADJ.DPEM.CD")
ind.df = wdi %>% filter(country.code %in% "USA") %>% filter(indicator.code %in% exp.var) %>% select(one_of(c("indicator.name","indicator.code")))
ind.df
```
Highest correlated variable
NY.ADJ.DPEM.CD - adjusted savings: particulate emission damage (current us$)
This indicator presents the damage caused by particulate emissions due to deaths.

Prediction and confidence:
```{r}
test.feat = test %>% select(-y)
pred = predict(model, newdata = test, interval = 'confidence')
pred = unscale(pred, miss.df)
pred.rname = rownames(pred)
pred.rname = as.numeric(pred.rname)
fin.df.row = final.df[pred.rname,] # filter the y's from final.df using the test row samples
# combine final.df data w/ predictions
test.pred = fin.df.row %>% select(one_of(c("country.name",y)))
test.pred = cbind(test.pred, pred)

head(test.pred, 10)
summary(test.pred)
```

Visualisation:
```{r}

########## GVIS MAP ##########

# 1. make a dataframe with country name and numbers
gdp.map = final.df %>% select(one_of(c("country.name", y)))
# filter out NAs just for the map
gdp.map$bool = is.na(subset(gdp.map, select = y))
gdp.map$bool = as.logical(gdp.map$bool)
gdp.map = gdp.map %>% filter(bool == FALSE) %>% select(-bool)
colnames(gdp.map) = c("country", "gdp.2015")
gdp.map$gdp.2015 = gdp.map$gdp.2015
gdp.map$scale = scale(gdp.map$gdp.2015) # standardization: (x - mean(x) / sd(x)
gdp.map$log = log(gdp.map$gdp.2015)

geo = gvisGeoChart(gdp.map, locationvar = "country",
                   colorvar = "gdp.2015",
                   options = list(
                     colorAxis = "{colors: ['burlywood', 'cornsilk', 'gold']}",
                     backgroundColor = "aliceblue",
                     height = 450
                   )
)

geo.stand = gvisGeoChart(gdp.map, locationvar = "country",
                            colorvar = "scale",
                            options = list(
                              colorAxis = "{colors: ['burlywood', 'cornsilk', 'gold']}",
                              backgroundColor = "aliceblue",
                              height = 450
                            )
)

geo.log = gvisGeoChart(gdp.map, locationvar = "country",
                   colorvar = "log",
                   options = list(
                     colorAxis = "{colors: ['burlywood','cornsilk', 'gold']}",
                     backgroundColor = "aliceblue",
                     height = 450
                   )
)

### CLEANING UP DATASETS FOR DISPLAY

########## SHINY ##########
# Shiny App
dashHead = dashboardHeader(title = "World GDP vs. Energy")

sideBar = dashboardSidebar(
  sidebarMenu(
    menuItem("2015 GDP Maps", tabName = "map"),
    menuItem("Data", tabName = "data"),
    menuItem("Trained Model", tabName = "model")
  )
)

body = dashboardBody(
  tabItems(
      tabItem(
        tabName = "map",
        fluidPage(column(12, align = "center",
                         h2("2015 GDP Map"),
                         htmlOutput("geo"),
                         h2("2015 Standardized GDP Map"),
                         htmlOutput("geo.stand"),
                         h2("2015 Log GDP Map"),
                         htmlOutput("geo.log")
                         )
                  )
        ),
      tabItem(
        tabName = "data",
        fluidPage(column(12, align = "center",
                         h2("2015 GDP and Energy Related Table"),
                            tabsetPanel(
                              tabPanel("Final DF", DT::dataTableOutput("final.df")),
                              tabPanel("Pred Df", DT::dataTableOutput("test.pred"))
                            
                            )
                          )
                  )
      ),
      tabItem(
        tabName = "model",
        fluidPage(column(12, align = "center",
                         h2("First Model Analysis"),
                         h4("Summary of the model using all of the explanatory variables"),
                         verbatimTextOutput(outputId = "sum.final.df"),
                         h4("Summary of the best fit model"),
                         verbatimTextOutput(outputId = "sum.test.pred")
                         )
                  )
        )
      )
  )


ui <- dashboardPage(dashHead,sideBar, body)

server = function(input, output) {
  
  output$geo <- renderGvis({
    map <- geo
  })
  output$geo.stand <- renderGvis({
    map <- geo.stand
  })
  output$geo.log <- renderGvis({
    map <- geo.log
  })
  output$final.df = DT::renderDataTable({
    final.df}, option = list(scrollX = TRUE))
  
  output$test.pred = DT::renderDataTable({
    test.pred}, option = list(scrollX = TRUE))
  
  output$sum.final.df <- renderPrint({
    fit = lm(NY.GDP.MKTP.CD ~ ., data = train)
    print(summary(fit))
  })
   output$sum.test.pred <- renderPrint({
    model = lm(NY.GDP.MKTP.CD ~ EG.CFT.ACCS.ZS + EG.ELC.ACCS.RU.ZS + NY.ADJ.DNGY.CD + NY.ADJ.DPEM.CD, data = train)
    print(summary(model))
  })

}

shinyApp(ui, server)
```

